\chapter*{Contributions}
\addchaptertocentry{Contributions}

In this section, the individual contributions for each chapter are stated in detail. 

\hfill \break
\noindent \textbf{Chapter 4 and 5:}

\noindent The original idea of DBM was developed by Marc Stieffenhofer, Michael Wand and Tristan Bereau. The implementation of DBM was carried out by Marc Stieffenhofer. The simulation data for syndiotactic polysteryne was provided by Chan Liu. Simulations of the octane and cumene systems were performed by Marc Stieffenhofer. Data analysis was carried out by Marc Stieffenhofer. The papers were written by Marc Stieffenhofer with critical commentary from Tristan Bereau and Michael Wand.

\hfill \break
\noindent \textit{Marc Stieffenhofer, Michael Wand, Tristan Bereau}\\
\textbf{Adversarial reverse mapping of equilibrated condensed-phase molecular structures}\\
Machine Learning: Science and Technology, Volume 1, Number 4\\
DOI: 10.1088/2632-2153/abb6d4\\
$\copyright$ IOP Publishing Ltd, 2020
\hfill \break

\noindent \textit{Marc Stieffenhofer, Tristan Bereau, Michael Wand}\\
\textbf{Adversarial reverse mapping of condensed-phase molecular structures: Chemical transferability}\\
APL Materials 9, Volume 9, Number 3\\
DOI: 10.1063/5.0039102\\
$\copyright$ AIP Publishing LLC, 2021
\hfill \break

\hfill \break
\noindent \textbf{Chapter 6:}

\noindent The original idea of the morphing approach was developed by Marc Stieffenhofer, Michael Wand and Tristan Bereau. Implementation of the approach was carried out by Marc Stieffenhofer. All simulations and data analysis were performed by Marc Stieffenhofer

\hfill \break
\noindent \textbf{Chapter 7:}

\noindent This chapter summarizes insights obtained during the course of a collaboration with Christoph Scherer, Falk May and Denis Andrienko. The original idea to use backmapping as a quality measure for coarse-grained models was developed by Marc Stieffenhofer and Denis Andrienko. Christoph Scherer parameterized the coarse-grained force field for Tris-Meta-Biphenyl-Triazine and carried out all simulations. Energy-based backmapping was performed by Christoph Scherer, while backmapping with DBM was performed by Marc Stieffenhofer. Data analysis and interpretation of the results were carried out by Marc Stieffenhofer.

\hfill \break
\noindent \textbf{Chapter 8:}

\noindent The work stems from a collaboration of Marc Stieffenhofer, Kirill Shmilovich, Moritz Hoffmann and Nick Charron. The project originates from the long program \textit{Machine Learning for Physics and the Physics of Learning} at the Institute for Pure \& Applied Mathematics that was held from 09.04.19 to 12.08.19 at the University of California, Los Angeles. The original idea of temporal coherent backmapping was introduced by Marc Stieffenhofer and further developed by all collaborators. Kirill Shmilovich carried out the implementation of the approach advised by Marc Stieffenhofer and Moritz Hoffmann. All-atom simulations of alanine dipeptide were performed by Kirill Shmillovich and all-atom trajectories for chignolin were provided by Jiang Wang. Coarse-grained force fields were parameterized by Nick Charron and coarse-grained simulations were performed by Nick Charron. Data analysis was performed by Kirill Shmillovich. The manuscript was written by Kirill Shmilovich, with critical commentary from all other collaborators.

\hfill \break
\noindent \textit{Kirill Shnilovich, Marc Stieffenhofer, Nick Charron, Moritz Hoffmann}\\
\textbf{Temporally coherent backmapping of molecular trajectories from coarse-grain to atomistic resolution}\\
arXiv preprint:2205.05213\\
DOI: 10.48550/ARXIV.2205.05213\\
$\copyright$ Creative Commons Attribution 4.0 International
